---
title: "Who were the Winners? \n\nWK6 SQL Q3"
author: "Sofia Ingersoll"
date: 2024-05-15
---

At the conclusion of the ASDN project the PIs decided to hand out first, second, and third prizes to the observers who measured the most eggs. Who won? Please use R and dbplyr to answer this question, and please submit your R code. Your code should print out:


```{r set_up, message = FALSE}
# let's load our libraries
library(tidyverse)
library(DBI)
library(dbplyr)
library(duckdb)
```
## Connect to Database
```{r db_connect}
# we are using 0.10 version of database, 0.9 will not work here
conn <- DBI::dbConnect(duckdb::duckdb(), dbdir = "database.db") 
```

#### View tables in DB
```{r view_tables}
# need to provide the connection u want to list
DBI::dbListTables(conn)
```
#### Select table in DB
```{r store_tables}
# pulling from db and storing as a table 
egg_table <- tbl(conn, "Bird_eggs")

observer_table <- tbl(conn, "Camp_Assignment")

personnel_table <- tbl(conn, "Personnel")

nest_table <- tbl(conn, "Bird_nests")

# quick peek at our tables
head(egg_table,3)

head(observer_table, 3)

head(personnel_table, 3)

head(nest_table,3)
```



This was a funky fun one to work through. Left, right, group, count, arrange & filter!
I defintely tried using a more SQL style of thinking my first few attempts and got some big total_egg numbers like 230!
```{r winners_table}
# let's look at our top observers!
top_winners <- egg_table %>% 
  # add all egg obs
  # `right join`
  # Vanessa Loverti	174			
  # Dylan Kessler	134			
  # Richard Lanctot	76
  left_join(nest_table, by = c("Book_page","Year", "Site", "Nest_ID")) %>% 
  # let's make our next join easy on us and match column names
  mutate(Abbreviation = Observer) %>% 
  # now we'll be able to see the full name of the observers
  right_join(personnel_table, by = "Abbreviation") %>% 
  # let's group by observer name
  group_by(Name) %>% 
  # and count how many eggs they weighed
  summarize(total_eggs = n()) %>% 
  # order them from greatest to least
  arrange(desc(total_eggs)) %>% 
  # we want the top observers!
  head(3)

top_winners  
```

#### Closing the connection
```{r}
DBI::dbDisconnect(conn, shutdown = TRUE)
```
